---
title: 'Discriminant and Classification Analysis of Prostate Cancer Prognostic Measurements'
subtitle: 'STAT 419: Multivariate Analysis — Spring 2026 Course Project'

author: 
  - Curtis Bradley
  - Ezli Cruz
  - Andrea Pareja
  - Andres Rocha Jayasinha

number-sections: true
number-depth: 2

format:
  html:
    css: styles.css
  pdf: 
    header-includes:
      - \usepackage{fvextra}
      - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
      - \renewcommand{\thesection}{\Alph{section}}
      - \renewcommand{\thesubsection}{\thesection.\arabic{subsection}}

crossref:
  sec-prefix: "Section"
---



```{r}
#| label: load-dataset
#| include: false
#| message: false
#| warning: false
required_pkgs <- c("MASS")
install_if_missing <- function(pkgs) {
  to_install <- pkgs[!pkgs %in% rownames(installed.packages())]
  if (length(to_install) > 0) {
    install.packages(to_install, repos = "https://cloud.r-project.org")
  }
}

install_if_missing(required_pkgs)
invisible(lapply(required_pkgs, library, character.only = TRUE))

source('_all_customized_functions.R') # custom multivariate functions (discrim, partial.F, etc.)

cancer <- read.csv('./data/cancer_419.csv', sep = ",", header = TRUE,
                   fileEncoding = "UTF-8-BOM")

# Encode the grouping variable as an ordered factor with readable labels
cancer$DIAGN <- factor(cancer$DIAGN,
                       levels = c(1, 2, 3),
                       labels = c("Low", "Moderate", "Severe"))

GROUP_VAR  <- "DIAGN"
QUANT_VARS <- setdiff(names(cancer), GROUP_VAR) # all predictor variables

set.seed(67)
```

# Introduction / Background {#sec-a}

A team of urologists set out to study how prostate-specific antigen (PSA) and a collection of clinical measurements relate to the progression of disease in men with advanced prostate cancer. The data set summarizes seven quantitative measurements taken on each of 97 such patients, along with a categorical variable indicating the severity of the disease. The goal of this project is to determine whether those seven measurements (or subset of them), considered jointly, can distinguish patients according to the severity of the seminal vesicle invasion, and to identify which measurements contribute most to that distinction.

The grouping variable for all analyses is **DIAGN**, which indicates the severity of the seminal vesicle invasion. This variable takes on three levels: low, moderate, and severe and is nearly balanced with `r sum(cancer$DIAGN == "Low")` low observations, `r sum(cancer$DIAGN == "Moderate")` moderate observations, and `r sum(cancer$DIAGN == "Severe")` severe observations.

The remaining seven variables are all quantitative and will serve as the predictors:

* **PSA** - serum prostate-specific antigen level, in mg/ml. A widely used biomarker for prostate disease; elevated values are associated with more advanced disease.
* **Volume** - volume of the prostate cancer, in cubic centimeters (cm3). A larger tumor volume generally signals a more advanced condition.
* **Weight** - weight of the prostate gland, in grams (g).
* **Age** - age of the patient, in years.
* **BPH** - amount of benign prostatic hyperplasia, in square centimeters (cm2). A measure of non-cancerous enlargement of the prostate.
* **Capsule** - degree of capsular penetration, i.e. the extent to which the tumor has pushed into or through the prostatic capsule.
* **Gleason** - a pathologically determined qualitative grade of the disease. Higher scores indicate a worse prognosis; in this sample the score takes only the integer values 6, 7, 8.

The analysis using these features went in three stages. First we examined the distribution of each predictor through histograms and summary statistics (@sec-explore). We then checked for multicollinearity amongst our features, and fitted linear discriminants to find the combination of variables that best separate the three severity groups while also testing these combinations for significance (@sec-discrim-importance). Finally, using the most influential variables, we constructed linear classification functions and applied them to classify patients, and evaluate how accurately the resulting rule determines the true severity groups (@sec-classification).


# Graphs and Summary Statistics {#sec-explore}

Before analysis, we will see the nature of each variable in our dataset. Below is a summary statistics table for each variable. 

```{r}
#| label: tbl-summary-stats
knitr::kable(data.frame(
  Variable = QUANT_VARS,
  Min      = sapply(cancer[QUANT_VARS], min, na.rm = TRUE),
  Q1       = sapply(cancer[QUANT_VARS], quantile, probs = 0.25, na.rm = TRUE),
  Mean     = sapply(cancer[QUANT_VARS], mean, na.rm = TRUE),
  Median   = sapply(cancer[QUANT_VARS], median, na.rm = TRUE),
  Q3       = sapply(cancer[QUANT_VARS], quantile, probs = 0.75, na.rm = TRUE),
  Max      = sapply(cancer[QUANT_VARS], max, na.rm = TRUE),
  SD       = sapply(cancer[QUANT_VARS], sd, na.rm = TRUE),
  row.names = NULL
), digits = 4)
```

Below we can also see histograms of each variable.


```{r}
#| label: fig-variable-histograms
#| fig-cap: "Histograms for all quantitative variables"
#| fig-width: 10
#| fig-height: 8

par(mfrow=c(3,3))

for (var in QUANT_VARS) {
  hist(cancer[[var]],
       main = paste("Histogram of", var),
       breaks = 15,
       xlab = var)
}

```


```{r}
#| label: define-transform-vars
#| include: false
# Variables that need transformation (identified from histograms)
TRANSFORMED_VARS <- c('PSA', 'Volume', 'Weight', 'BPH', 'Capsule')
LOG_VARS         <- c('PSA', 'Volume', 'Weight')         # log(x) — no zeros
LOG1P_VARS       <- setdiff(TRANSFORMED_VARS, LOG_VARS)  # log(x+1) — contain zeros

# Pre-build formatted strings for use in inline R prose
fmt <- function(v) paste0("`", v, "`")
TRANSFORMED_VARS_fmt <- paste(fmt(TRANSFORMED_VARS), collapse = ", ")
LOG_VARS_fmt         <- paste(fmt(LOG_VARS), collapse = ", ")
LOG1P_VARS_fmt       <- paste(fmt(LOG1P_VARS), collapse = " and ")
KEPT_VARS_fmt        <- paste(fmt(setdiff(QUANT_VARS, TRANSFORMED_VARS)), collapse = " and ")
LOG_VARS_log_fmt     <- paste0("log(", LOG_VARS, ")")
LOG_VARS_log_str     <- paste(fmt(LOG_VARS_log_fmt), collapse = ", ")
```

Based on the histogram it looks like the variables `r TRANSFORMED_VARS_fmt` are non normal.  We can try to apply transformations on them to see if we can make them look more normal.
```{r}
#| label: fig-transformed-histograms
#| fig-cap: "Histograms for all transformed variables"
#| fig-width: 10
#| fig-height: 8

par(mfrow=c(2,3))

# Variables requiring transformation (identified from raw histograms above)
TRANSFORMED_VARS <- c('PSA', 'Volume', 'Weight', 'BPH', 'Capsule')
LOG_VARS <- c('PSA', 'Volume', 'Weight')          # strictly positive
LOG1P_VARS <- setdiff(TRANSFORMED_VARS, LOG_VARS) # contain zeros 

for (var in LOG_VARS) {
  hist(log(cancer[var][,]), main=paste("Histogram of LOG(", var, ")", sep=''), breaks=15, xlab=paste('log(',var,')',sep=''))
}

for (var in LOG1P_VARS) {
  hist(log1p(cancer[var][,]), main=paste("Histogram of LOG(", var, " + 1)", sep=''), breaks=15, xlab=paste('log(',var,' + 1)',sep=''))
}

plot.new()  # pad the 2x3 grid (5 plots, 1 empty cell)
```

:::{.callout-note}
`BPH` and `Capsule` are transformed with a log + 1 transformation because they contain 0 values. See @tbl-summary-stats
:::

Looking at @fig-transformed-histograms, `r LOG_VARS_fmt` all appear to be roughly normal, however `r LOG1P_VARS_fmt` do not. Because Discriminant Analysis has a normality assumption we drop them from our data and only use `r LOG_VARS_log_str`, and the untransformed `r KEPT_VARS_fmt` for our analysis.


Below is a scatterplot matrix of the transformed dataset.
```{r}
#| label: fig-scatterplot-matrix
#| fig-cap: "Scatterplot matrix of transformed dataset"
#| fig-width: 10
#| fig-height: 8

# Build the modelling dataset 
cancer.mod <- cbind(cancer[setdiff(colnames(cancer), TRANSFORMED_VARS)], log(cancer[LOG_VARS]))

# Color points by DIAGN group; exclude DIAGN column from the numeric panels
pairs(cancer.mod[, -1],
      main = "Scatterplot Matrix of Quantitative Variables",
      pch = 19, cex = 0.5,
      col = c("steelblue", "darkorange", "firebrick")[cancer.mod$DIAGN])
```


## Outlier Identification {#sec-outlier-id}

After looking at the data we can see some trends and possibly some outliers. Of note there is a point that appears to have a higher weight than the others. We can also check for outliers by performing a slightly more statistical method. Assuming that we have an MVN already, we can see how far away each dot is from the center, and the probability of seeing an observation that extreme.
```{r}
#| label: tbl-outlier-identification
# Use mahalanobis to measure distance from centroid
md <- mahalanobis(
  x      = cancer.mod[, -1],
  center = colMeans(cancer.mod[, -1], na.rm = TRUE),
  cov    = cov(cancer.mod[, -1], use = "complete.obs")
)

# Bonferroni correction
sig_level <- 0.05 / nrow(cancer[, -1])

# Upper-tail p-value: probability of a chi-sq value this large under MVN
# (large distance -> small p -> flagged as outlier)
observation_p <- pchisq(md, df = ncol(cancer.mod[, -1]), lower.tail = FALSE)

knitr::kable(cbind(cancer, observation_p)[which(observation_p < sig_level), ])
```

Based on the results we can see that 41 appears to be different from the dataset. Looking at the output of @tbl-outlier-identification and comparing it to @tbl-summary-stats, all of the variables seem to be within normal range, with the exception of `Weight`. The weight of `r cancer$Weight[41]` grams is significantly higher than the normal range. In fact, it is `r round((cancer$Weight[41] - mean(cancer$Weight)) /  sd(cancer$Weight),3)` standard deviations away from the mean weight. For this reason we excluded this datapoint from our analysis.


```{r}
#| label: drop-outlier
cancer.mod <- cancer.mod[-41,]
```


We then checked our MVN assumptions (see @sec-mvn-assumption).

## MVN Assumptions {#sec-mvn-assumption}

```{r}
#| label: fig-mvn-plots
#| fig-cap: "MVN Q-Q Plots"
#| fig-width: 10
#| fig-height: 8

par(mfrow=c(2,2))

check.mvnorm.plot(cancer.mod[,-1])
mtext("All Values",side=3)
check.mvnorm.plot(cancer.mod[cancer.mod$DIAGN == "Low",][,-1])
mtext("DIAGN=Low",side=3)
check.mvnorm.plot(cancer.mod[cancer.mod$DIAGN == "Moderate",][,-1])
mtext("DIAGN=Moderate",side=3)
check.mvnorm.plot(cancer.mod[cancer.mod$DIAGN == "Severe",][,-1])
mtext("DIAGN=Severe", side=3)

```


These plots looks pretty much like they're all MVN. The last check is to make sure they all have the same covariance matricies. To do this we use Box's M test.

```{r}
#| label: tbl-box-m
# Box's M tests H0: all groups share a common covariance matrix (required for LDA)
cov_mats <- list(
  Low      = cov(cancer.mod[cancer.mod$DIAGN == "Low",      -1], use = "complete.obs"),
  Moderate = cov(cancer.mod[cancer.mod$DIAGN == "Moderate", -1], use = "complete.obs"),
  Severe   = cov(cancer.mod[cancer.mod$DIAGN == "Severe",   -1], use = "complete.obs")
)

n_obs <- unname(as.vector(table(cancer.mod$DIAGN)[c("Low", "Moderate", "Severe")]))

Box.M(cov_mats, n_obs)
```

:::{.callout-caution}
We actually failed Box's M test. However we still decided to proceed with discriminant analysis as it looks close enough.
:::

Based on the output it appears all groups follow a multivariate normal distribution. This allows us to move on to discriminant analysis.


# Discriminant and Classification Analysis {#sec-discrim-class}

## Correlated Quantitative Variables (Multicollinearity) {#sec-collinearity}
```{r}
#| label: tbl-corr
knitr::kable(cor(cancer.mod[,-1]))
```

To assess for multicollinearity, we analyze the correlation coefficients for each pair of quantitative variables. Our scatterplot matrix shows that there is no prominent linear relationship between any two variables. We can then look at the calculated correlation coefficients table. Using a correlation threshold of $|r| \ge 0.80$, none of our variables have high correlation. We will not flag any of our variables for multicollinearity. We will keep all of our `r length(cancer.mod) - 1` quantitative variables. 


## Discriminant Analysis {#sec-discrim-importance}

### Discriminant Functions and Variable Importance

```{r}
#| label: tbl-discrim-vals
da <- discrim(cancer.mod[,-1], cancer.mod[,1])

n_ld     <- ncol(da$a.stand)
std_coef <- da$a.stand
rownames(std_coef) <- colnames(cancer.mod[,-1])
colnames(std_coef) <- paste0("LD", seq_len(n_ld))  # label columns LD1, LD2, ...

knitr::kable(std_coef)
```

```{r}
#| label: lda-formatted-output
#| include: false
lda_latex <- function(coefs, log_transformations = NULL, digits = 4,
                      names = colnames(coefs)) {
  coefs <- as.data.frame(coefs)
  vars  <- rownames(coefs)

  # Fallback column names if none supplied
  if (is.null(names)) names <- paste0("LD", seq_along(coefs))

  # Wrap logged variables in \log(...)
  vars_latex <- ifelse(
    vars %in% log_transformations,
    paste0("\\log(", vars, ")"),
    vars
  )

  equations <- vapply(seq_along(coefs), function(j) {
    terms <- paste0(
      sprintf(paste0("%.", digits, "f"), coefs[[j]]),
      "\\,", vars_latex,
      collapse = " + "
    )
    terms <- gsub("\\+ -", "- ", terms)
    paste0(names[j], " &= ", terms)   # & aligns the = signs, removes leading indent
  }, character(1))

  paste0(
    "$$\\begin{aligned}\n",
    paste(equations, collapse = " \\\\\n"),
    "\\end{aligned}$$"                  # no trailing \n -> no extra space below
  )
}
```

A linear discriminant analysis was performed to observe possible separation between the groups of severity of seminal vesicle invasion and what variables contribute the most in that separation. The discriminant analysis resulted in 2 discriminant functions: `r lda_latex(std_coef, log_transformations = LOG_VARS)`

When looking at the importance of each variable in the first discriminant function, we get:

```{r}
#| label: tbl-ld1-coef
# Sort variables by absolute LD1 coefficient (largest = most important)
ld1_order <- names(sort(abs(std_coef[, 1]), decreasing = TRUE))

knitr::kable(
  data.frame(
    Variable = ld1_order,
    LD1 = std_coef[ld1_order, 1],
    row.names = NULL
  ),
  digits = 4,
)
```



The table of the standardized coefficients allows us to see that `r ld1_order[1]` seems to strongly contribute in group separation of severity of seminal vesicle invasion. Then `r ld1_order[2]` seems to have moderate contributions while `r paste(ld1_order[3:length(ld1_order)], collapse = ", ")` seem to have the weakest contribution.


### Test of Significance of the Discriminant Functions

To see which of linear discriminant functions are significant in group separation, we will test:


__(1)__
$H_0: \alpha_1 = \alpha_2 = 0$,
$H_a: \text{at least one } \alpha \ne 0$

__(2)__
$H_0: \alpha_2 = 0$,
$H_a: \alpha_2 \ne 0$

Since we have two discriminant functions, there will be 2 tests performed (listed above), we must adjust our level of significance to $\frac{0.05}{2} = 0.025$. The results are:

```{r}
#| label: tbl-discr-sig
discr_sig_tbl <- discr.sig(cancer.mod[,-1], cancer.mod[,1])
knitr::kable(discr_sig_tbl)
```

LD1: With a p-value of `r sprintf("%.4f", discr_sig_tbl["LD1", "p.values"])` being less than our adjusted significance level of 0.025, we have enough evidence that the first linear discriminant function works well in group separation of severity of seminal vesicle invasion. 

LD2: With a p-value of `r sprintf("%.4f", discr_sig_tbl["LD2", "p.values"])` being larger than our adjusted significance level of 0.025, we don't have enough evidence that the second discriminant function is useful in group separation of severity of seminal vesicle invasion.

### Significance of Each Variable, Adjusted for the Others

After seeing the standardized coefficients of the first discriminant function, the variable PSA might be important in group separation. Here are the formal partial F tests for each variable: 

$H_0$: Variable (y) does not contribute to group separation of severity of seminal vesicle invasion, adjusting for the other variables.

$H_a$: Variable (y) does contribute to group separation of severity of seminal vesicle invasion, adjusting for the other variables.


```{r}
#| label: tbl-discr-partial-f
partial_f_tbl   <- partial.F(cancer.mod[,-1], cancer.mod[,1])
n_partial_tests <- nrow(partial_f_tbl)  # one test per variable
```

Since there are `r n_partial_tests` tests being performed, we must adjust our significance level to $\alpha^* = \frac{.05}{`r n_partial_tests`} = `r round(.05/n_partial_tests,4)`$.

```{r}
#| label: tbl-discr-partial-f-display
knitr::kable(partial_f_tbl)
```

From our results, PSA significantly contributes the most in group separation of severity of seminal vesicle invasion, adjusting for the other variables. 

All other variables do not individually contribute significantly in group separation of severity of seminal vesicle invasion, after adjusting for the other variables.

### Plot of the First Two Discriminant Functions

Below is a plot of the two linear discriminant functions.

```{r}
#| label: fig-discr-plot
#| fig-cap: "Discriminant Plots"
#| fig-width: 10
#| fig-height: 8

discr.plot(cancer.mod[,-1], cancer.mod[,1], title='Discriminant Plot for Diagnosis Groups', leg = levels(cancer.mod[,1]))

```

From this plot, it is visually evident the first discriminant function (LD1) captures the group separation well of the three groups of low, moderate, and severe seminal vesicle invasion. The severe severity of seminal vesicle invasion is on the left while the moderate level is on the right. Low severity of seminal vesicle invasion seems grouped in between the moderate and severe levels. 

The second discriminant function (LD2) does not capture the separation nature of the data as projections of the points on the line are clumped together. This further proves that LD1 is the better discriminant function that works well in group separation. 


## Classification Analysis {#sec-classification}

For classification analysis we first select the top 4 most important variables from the LD1 importance ranking.

```{r}
#| label: classification-setup
#| include: false
TOP_K    <- 4
top_vars <- head(ld1_order, TOP_K)  # top 4 variables by LD1 importance
```

Looking at @tbl-ld1-coef, the top four variables, in order of importance, are `r paste(top_vars, collapse = ", ")`.

### Linear Classification Functions

The linear classification function for each severity group is given below. An observation is assigned to the group whose function returns the largest score.

```{r}
#| label: tbl-classification-functions
lc <- lin.class(cancer.mod[, top_vars], cancer.mod[, 1])

# Reshape coefficients: rows = variables, columns = groups
coefs_byvar <- t(lc$coefs)
rownames(coefs_byvar) <- top_vars
colnames(coefs_byvar) <- levels(cancer.mod[,1])

# Prepend the constant term (c0) as its own row
lcf_table <- rbind(Constant = lc$c.0, coefs_byvar)
colnames(lcf_table) <- levels(cancer.mod[,1])

knitr::kable(lcf_table, digits = 4)
```

```{r}
#| label: lcf-formatted-output
#| include: false
lcf_latex <- function(lcf, log_transformations = NULL, digits = 4) {
  vars       <- rownames(lcf)[-1]                     # drop the Constant row
  consts     <- lcf["Constant", ]
  vars_latex <- ifelse(vars %in% log_transformations,
                       paste0("\\log(", vars, ")"), vars)

  equations <- vapply(seq_along(colnames(lcf)), function(j) {
    terms <- paste0(
      sprintf(paste0("%.", digits, "f"), lcf[-1, j]),
      "\\,", vars_latex, collapse = " + "
    )
    const <- sprintf(paste0("%+.", digits, "f"), consts[j])
    eq    <- paste0("L_{", j, "}(y) = ", terms, " ", const)
    gsub("\\+ -", "- ", eq)
  }, character(1))

  paste0("$$\\begin{aligned}\n",
         paste0(gsub(" = ", " &= ", equations), collapse = " \\\\\n"),
         "\\end{aligned}$$")
}
```

Written out explicitly, where 1 = Low, 2 = Moderate, and 3 = Severe: `r lcf_latex(lcf_table, log_transformations = LOG_VARS)`

### Classifying Observation #1

We apply the three functions to the first observation, compute each score, and assign the group with the largest value.

```{r}
#| label: tbl-classify-obs1
groups <- levels(cancer.mod[,1])

# Compute LCF score for a single observation against each group
lcf_scores <- function(x_vec) {
  sapply(groups, function(g) lc$c.0[match(g, groups)] + sum(coefs_byvar[, g] * x_vec))
}

obs1_x      <- cancer.mod[1, top_vars]         # predictor values for observation 1
obs1_scores <- lcf_scores(obs1_x)              # score against each group's function
obs1_pred   <- groups[which.max(obs1_scores)]  # predicted group = highest score
obs1_actual <- as.character(cancer.mod[1, 1])  # true group label

knitr::kable(
  data.frame(Group = groups, Score = round(obs1_scores, 4), row.names = NULL),
  digits = 4,
)
```

The largest score belongs to the **`r obs1_pred`** group, so Observation #1 is classified as `r obs1_pred`. Its actual group is `r obs1_actual`, so this prediction is `r if (obs1_pred == obs1_actual) "correct" else "incorrect"`.

### Confusion Matrix and Error Rates

Using resubstitution (apparent) classification on all observations:

```{r}
#| label: tbl-confusion-matrix
cls       <- rates(cancer.mod[, top_vars], cancer.mod[,1], method = "l")
confusion <- cls[["Confusion Matrix"]]
ACCR      <- cls[["Correct Class Rate"]]
APER      <- cls[["Error Rate"]]

knitr::kable(as.data.frame.matrix(confusion))
```

```{r}
#| label: tbl-error-rates
n         <- nrow(cancer.mod)
n_correct <- round(ACCR * n)  

knitr::kable(
  data.frame(
    Metric = c("Apparent Correct Classification Rate (ACCR)",
               "Apparent Error Rate (APER)", "N", "Correct"),
    Value  = c(ACCR, APER, n, n_correct)
  ),
  digits = 4
)
```

We have an overall correct classification rate of `r round(ACCR * 100, 1)`%, indicating a good discrimination model. The `r round(APER * 100, 1)`% of misclassifications fall between adjacent severity groups; there are no misclassifications between the Low and Severe groups. With a high correct classification rate, the model provides accurate prediction of the severity groups.


# Summary {#sec-summary}

The goal of this project was to determine whether `r length(QUANT_VARS)` clinical measurements could effectively distinguish prostate cancer patients according to the severity of seminal vesicle invasion. Based on our analysis, the answer is yes. Among the measured variables, `r ld1_order[1]` was the most influential in separating the groups, as evidenced by its significance in the discriminant analysis (@sec-discrim-importance) and its large contribution to the first linear discriminant function. In addition, plotting the observations with respect to LD1 and LD2 (@fig-discr-plot) revealed strong separation among the severity groups, particularly along LD1.

This conclusion is further supported by the classification results (@sec-classification). Using the linear classification functions, the apparent correct classification rate (ACCR) was `r round(ACCR, 4)`, meaning that `r n_correct` of the `r n` observations were correctly classified. The confusion matrix (@tbl-confusion-matrix) also indicates that the model avoided the most severe types of misclassification: no patients in the low-severity group were classified as severe, and no patients in the severe group were classified as low. Instead, the misclassifications were limited to adjacent severity categories, such as low-severity observations being classified as moderate. Overall, these results suggest that the clinical measurements, particularly `r ld1_order[1]`, provide meaningful discrimination among levels of seminal vesicle invasion severity.

One thing in this analysis did stand out, mainly that no variables were collinear enough to drop in @sec-collinearity, even though we expected PSA, Volume, and Capsule might have been redundant; their correlations were moderate, but never reached our thresholds.

If we were to repeat this project, we would make two key improvements. First, we would explore alternative data transformations, either in place of or in combination with log transformations, such as the Box–Cox transformation. This would be particularly useful for variables such as BPH and Capsule, which remain skewed even after applying a log transformation. Second, we would implement leave-one-out cross-validation to estimate the classification error rate. This would allow us to assess whether the selected transformations and resulting classification model generalize well to new observations, rather than overfitting to the particular dataset used in this analysis.

Finally, one variable we wish the data would have included is a direct measure of tumor stage (such as a TNM clinical stage) or a PSA reading taken over time rather than at a single point. More generally some variable(s) that would give information about the trajectory of the disease for a given patient would help in getting cleaner separation for the observations on the border of moderate and severe, which were the toughest to classify with measurements available.
